Framing active school travel in Ontario: , or how spinach is good for you

### 2023-05-23 editorial-materials/YTRA-D-22-01243-response-to-reviewers.Rmd

---
title: "How is active school travel framed in Ontario, Canada?"
output: pdf_document
link-citations: yes
linkcolor: blue
bibliography: 
  - '`r system("kpsewhich ../AST-Framing-Ontario/bibliography.bib", intern=TRUE)`'
---

# Responses to reviewers

We wish to thank both reviewers for their constructive feedback. We feel that the paper has been much improved as a result of responding to their comments.

## Reviewer 1

### Comment 1 {#r1-comment-1}

> The authors have used topic modelling to study policy centred and academic writing about active transportation to school in Ontario, Canada. I was quite fascinated by the approach - the paper presents a new way of looking at a vast grey and peer reviewed literature on school travel that has developed over several decades.

Thank you for your kind comments and generous feedback.

### Comment 2 {#r1-comment-2}

> I think the authors should offer more detail on their search strategy beyond, in section 4.1.2, simply saying - we conducted a search using Web of Science. Digging into and report search method - and why/how such decisions are made is an important piece of work like this. Why did you limit your search to the fields indicated? How did you arrive at your search strings/keywords? Having some reporting here, similar to what one would find in a review paper (systematic, scoping, or narrative for that matter) would be useful.

Thank you for this comment. We were too terse with our reporting of the search protocol in the earlier version of the paper. This has been rectified in the current version, where we add relevant details, as follows.

The search was conducted using Web of Science's Core Collection. This was done in the Winter of 2021 and the parameters we used were {active OR walk* or cycl* or bicycl*} AND {"school travel"}. Initially, we found 322 papers, which we reduced to 250 after selecting only papers in the fields of transportation, planning, urban studies, geography, and public health. These fields were chosen because they had the greatest volume of documents and/or were of disciplinary interest to the authors. This list with 250 documents was then manually curated by the authors to ensure that all documents were relevant. This was done by checking the title and reading the abstract of the papers. For example, a paper with the title "Impact of automated photo enforcement of vehicle speed in school zones" [@quistberg2019impact] was excluded as being tangential to our research; the abstract of another paper revealed that it was mainly concerned with survey methods: "Common methods for measuring mode share include Hands Up surveys and family surveys, but these require teacher and parental involvement." [@sersli2019getting]. After this process, our academic corpus comprised 227 journal articles that were readied for analysis.

### Comment 3 {#r1-comment-3}

> I think section 5, and the paper overall, could be strengthened by writing a bit more on the relationship between this automated topic modelling approach, and traditional forms of qualitative thematic analysis - some reflection on the possibility of generating similar or different results, using either approach - would be quite fascinating and helpful. I couldn't help but wonder if the authors considered this approach to be a replace for .... or in addition too, traditional - less automated approaches. Certainly one advantage is just being able to handle the volume of data.

Thank you for this comment. We have added the following to the concluding section:

Finally, we would like to close with a brief note on the methodology. Natural language processing and analysis of frames are still not widely adopted in transportation research. Based on our experience, we tend to see computer-assisted text analysis as a complement rather than a substitute to expert reading of the documents. We would not recommend to proceed to interpretation without reading the documents, but the supervised automation helps to detect patterns that can be validated by an expert based on spot checks or more in-depth reading of the original sources. Like with any qualitative research, interpretation of the frames may differ by expert; in this research we aimed to develop a consensus interpretation by involving all authors in the discussion of the results. Furthermore, the research was conducted as an open, reproducible project: all code and data are available in a public repository^[\url{https://github.com/desjae/AST-Framing-Analysis-Ontario}], which means that all our assumptions are open to scrutiny. Moreover, other researchers can refer to the repository to expand or challenge our conclusions.

### Comment 4 {#r1-comment-4}

> This brings me to my primary concern - I couldn't help but wonder if this piece might be a better fit to TR-B or C - given the emphasis on demonstrating method/tools - and the production of conclusions that are rather unsurprising - the last sentence of the paper alone is something that has been the focus of much work, in practice, for decades.

Thank you for sharing this concern. The methods are not really new outside of their application to this research. Natural language processing is a field mature enough to have its own textbooks; framing analysis is also widespread, although its use in transportation research remains limited. We do not think of either method as emerging technologies or sufficiently innovative to make this a "methodologica" paper. Instead, what we think is new is the ability of these methods to reveal something interesting about the effectiveness of transportation policy.

The last sentence in the original paper reads "STP stakeholders should emphasize why AST rates should increase in local communities and how the negative effects of nonactive modes to school may impact children's health and wellbeing". It is true that there is abundant research about the negative effects of nonactive modes, and most experts would agree that those negative effects explain why active travel rates should increase. What is new is what you astutely note in the following comment: despite all this wealth of knowledge, we don't seem to be further along in terms of convincing the public of buying into active travel policies. Research into how active travel is framed can help us understand what needs to be done to craft more effective communications and shared meanings of the relevant issues.

### Comment 5 {#r1-comment-5}

> The same can be said of the first sentence of section 7.2 - i.e, the problem is the car, and so we need to make cars seem less helpful to the cause. At the same time, the wide ranging scope of the analysis somewhat interestingly tells us that despite all of this research and policy work - we are somehow no further along (overall) than perhaps we were at the very beginning.

This is a very shrewd comment that in our opinion bangs the nail right on the head: the problem is not so much that we don't know that the problem is the car. The problem is that we don't know how to effectively communicate that knowledge to a variety of stakeholders for the purpose of creating shared meanings that can support policy development and implementation. By examining the case of AST in Ontario we find that the framing of transportation in communities is reasonable but anodyne. It downloads responsibility for change to individuals and does not make an effective call for collective action. The tone is positive and does not convey urgency about some critical stakes (e.g., climate change). In essence, the framing can be likened to "eating spinach is good for you; if you decide to eat spinach we can tell you where to buy some". We cannot imagine anyone arguing that eating spinach is bad, but it is questionable how effective this message would be to persuade people to eat spinach.

We tried to capture these ideas in the rewritten abstract of the paper:

Active school travel (AST) is a relevant policy issue in Ontario, Canada, as reflected in provincial support for school travel planning (STP) efforts. Two pillars of school travel plans are education and encouragement. For these pillars to stand, all relevant stakeholders must be cognizant of and understand well the issues and stakes. For this reason, how AST is communicated to them is essential. In this research we adopt frame analysis to investigate the way various organizations communicate around AST to the public. A frame is a central organizing idea or story line that provides meaning to a particular phenomenon, and helps to set the parameters for conversations about policy. In the case of AST, framing can influence what policy alternatives are perceived as available by children, parents, and their wider communities. To investigate framing of AST in Ontario we use natural language processing techniques to analyze publicly available documents from Ontario stakeholders involved in school travel planning. We then compared the findings from these documents to a selection of academic studies on AST. We conclude that framing of AST in Ontario is mostly empirical-scientific and largely in agreement with academic research on AST. On the other hand, it it does not appeal to moral reasons for change, and is therefore oddly soothing and lacking in urgency. Furthermore, the frames used download responsibility for change to households, which limits the scope of policy alternatives by keeping collective action out of the frame.

\newpage

## Reviewer 2

### Comment 1 {#r2-comment-1}

> This paper is well written and the analysis is conducted rigorously and it is interesting.

We are grateful for your comments and feedback.

### Comment 2 {#r2-comment-2}

> However in my opinion is not within the scope of Transportation Research Part A. I did not understand the link with the policy and practice.

We wish to reiterate our response to [Reviewer 1](#r1-comment-4). We part from the working hypothesis that if the Editor, having perused the paper, decided to send it out for peer-review, they must have decided that the paper is within the scope of the journal.

Having said that, we made a renewed effort to clarify the link of our research to policy and practice. In particular, we think Reviewer 1 right on target with their [comment](#r1-comment-5): despite the mass of academic research in transportation, we do not seem to be further along in effectively communicating with the public the ever higher stakes of our ongoing love affair with the car. The problem does not seem to be the lack of policies, but awareness and buy-in by the public. Why is this the case when it comes to active school travel?

The paper has been reworked to (hopefully) improve the link between framing and policy. The argument goes as follows:

1. "Public policies are directly tied to the constant shaping of collective identities" [@beland2014developing, p. 551; also see beland2008nationalism; @akerlof2010identity]

2. Collective identities (and therefore policy) are shaped by frames: frames can activate or restrict particular responses in the intended audience [@panFramingAnalysisApproach1993]. They do this by position existing solutions as suitable to issues, which limits awareness of other policy approaches [@mahFramecriticalPolicyAnalysis2014] or, contrariwise, by revealing more diverse policy options than conventional wisdom would have [@bosomworth2015climate].

3. Content analysis of documents in the case of Ontario uncovers some reasons why AST, despite being a policy issue, has struggled to reverse motorization trends.

4. Active school travel is framed in Ontario in such a way that it downloads the responsibility for action to individuals; it does not make a convincing call for collective action. It is also fairly neutral at a time when some stakes (e.g., climate change) are extremely high. By adopting an empirical-scientific approach, it does not appeal to moral reasons for change. And it ends up being lukewarm and anodyne.

We tried to capture these ideas in the rewritten abstract of the paper:

Active school travel (AST) is a relevant policy issue in Ontario, Canada, as reflected in provincial support for school travel planning (STP) efforts. Two pillars of school travel plans are education and encouragement. For these pillars to stand, all relevant stakeholders must be cognizant of and understand well the issues and stakes. For this reason, how AST is communicated to them is essential. In this research we adopt frame analysis to investigate the way various organizations communicate around AST to the public. A frame is a central organizing idea or story line that provides meaning to a particular phenomenon, and helps to set the parameters for conversations about policy. In the case of AST, framing can influence what policy alternatives are perceived as available by children, parents, and their wider communities. To investigate framing of AST in Ontario we use natural language processing techniques to analyze publicly available documents from Ontario stakeholders involved in school travel planning. We then compared the findings from these documents to a selection of academic studies on AST. We conclude that framing of AST in Ontario is mostly empirical-scientific and largely in agreement with academic research on AST. On the other hand, it it does not appeal to moral reasons for change, and is therefore oddly soothing and lacking in urgency. Furthermore, the frames used download responsibility for change to households, which limits the scope of policy alternatives by keeping collective action out of the frame.

### Comment 3 {#r2-comment-3}

> The research finds that a specific policy is not working as intended and it seems to be very expensive. How can this result be generalised and used to say if any policy is "good value for money"? The authors need to convince me about the practical application of their method. Can you identify clear policy recommendations that can be used in practice?

Point 4 [above](#r2-comment-2) highlights some things that we learn from this research, and suggests some ways to change the way AST is framed to make them more efficient.

### Comment 4 {#r2-comment-4}

> Otherwise the paper, even if interesting, is only methodological and would fit better some journals that focuses on text analysis.

Thank you for your opinion. We will defer to the editor on the matter of scope of the journal.

\newpage

## References

### 2023-05-23 AST-Framing-Ontario/AST-Framing-Ontario.Rmd
Revisions after feedback from TR-A

---
title: How is active school travel framed in Ontario, Canada?
author:
  - name: Elise Desjardins
    email: desjae@mcmaster.ca
    affiliation: McMaster University
  - name: Jason Lam
    email: lamj54@mcmaster.ca
    affiliation: McMaster University
  - name: Darcy Reynard 
    email: reynard@ualberta.ca
    affiliation: University of Alberta
  - name: Damian Collins
    email: damian.collins@ualberta.ca
    affiliation: University of Alberta
  - name: E. Owen D. Waygood
    email: owen.waygood@polymtl.ca
    affiliation: Polytechnique Montreal
  - name: Antonio Paez
    email: paezha@mcmaster.ca
    affiliation: McMaster University
    correspondingauthor: true
    footnote: 1
address:
  - code: McMaster University
    address: School of Earth, Environment and Society, 1280 Main St. W, Hamilton, Ontario, Canada L8S 4K1
  - code: University of Alberta
    address: Department of Earth and Atmospheric Sciences, 1-26 Earth Sciences Building, Edmonton, Alberta, Canada T6G 2E3
  - code: Polytechnique Montreal
    address: Department of Civil, Geological and Mining Engineering, 2500, chemin de Polytechnique Montreal, Montreal, Quebec, Canada H3T 1J4
footnote:
  - code: 1
    text: "Corresponding author."
abstract: |
  Active school travel (AST) has been increasingly encouraged by various stakeholders in Ontario, Canada through efforts such as school travel planning. Education strategies like workshops or resources that promote AST are commonly implemented. The framing of AST through such strategies may influence how walking and bicycling to school are perceived by parents. It may also draw attention to AST as an issue affecting children's health which could motivate behaviour change. We used natural language processing, including topic modelling, to examine how AST is framed in publicly available documents from Ontario stakeholders involved in school travel planning. We then compared the findings from these documents to a selection of studies on AST and explored similarities between the two. We found that AST is framed in two ways: i) as a health and environmental issue; and ii) as an accessible and feasible transport option for children and parents. The frames encourage children and parents to adopt AST given its health and environmental benefits by providing resources to support behaviour change. The benefits of AST and strategies to support AST that are communicated by stakeholders are consistent with the evidence from academic research. While these frames present AST in a positive light, they may not encourage parents to view current household travel behaviours as unhealthy for their children or their community. Stakeholders promoting AST in Ontario should further problematize the decline of AST and challenge the norm of driving children to school.
keywords: 
  - active school travel
  - school travel planning
  - natural language processing
  - framing analysis
journal: "Transportation Research Part A: Policy and Practice"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: bibliography.bib
linenumbers: false
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: natbib
---

@@

Walking and bicycling to school, commonly known as active school travel (AST), have declined in North America for decades [@rothmanDeclineActiveSchool2018], but the appetite for these modes is beginning to rise in many Canadian communities. For instance, research in Toronto, Ontario found that 40% of children who were driven to school would like to travel by bicycle instead [@laroucheRatherBikeSchool2016]. Another study in London, Ontario, reported similar findings with respect to children's preference for active travel [@larsenRouteBasedAnalysisCapture2012]. From a public health perspective, increasing active travel to school represents a major opportunity to improve children's health and wellbeing.

To promote this trend, in 2017 the Government of Ontario began to support communities develop AST initiatives through a program called _Ontario Active School Travel_ (OAST). By 2021, the program, run by Green Communities Canada had awarded over CAD 2 million and provided resources to over 25 projects across Ontario [@GreenCommunities2016]. A popular intervention in many communities is school travel planning (STP) [see @GreenCommunitiesProjects]. In this context, STPs are "school-specific" interventions led by a facilitator who works with a committee of stakeholders from diverse sectors including education, planning, transportation, and public health to develop action plans [@buliungSchoolTravelPlanning2011; @mammenSchoolTravelPlanning2014]. The five-step process involves identifying barriers to AST and implementing solutions or activities to make AST safer and more convenient: 1) setup of the program; 2) data collection and problem identification; 3) action planning; 4) implementation; 5) monitoring and evaluation [@buliungSchoolTravelPlanning2011; @langUnderstandingModalChoice2011].

Framing AST by STP stakeholders to a target audience (e.g., parents or the general public) can raise awareness of the relevant issues and influence how active travel to school is perceived. Gamson and Modigliani [-@gamsonChangingCulture1987] define a frame as a “central organizing idea or story line that provides meaning” to a particular phenomenon. A frame can enable individuals "to locate, perceive, identify, and label" information pertaining to various dimensions of an issue [@goffmanFrameAnalysisEssay1974]. Therefore, stakeholders involved in STP efforts can play a role in shaping public perception about AST to attract greater attention and recognition as a "problem" that needs to be addressed through behaviour change or new policies.

After significant investment of human and financial resources to boost rates of AST in Ontario over the past few years, in this paper we ask this central question: How is AST framed to the public in Ontario? The following questions are subsidiary and help us address this main question: What benefits of AST are communicated to the public? What solutions are proposed to increase AST?

We approach this question by means of natural language processing, and examine how three STP stakeholder groups in Ontario (municipalities, school boards, and transportation consortia) frame AST. We assemble a corpus of texts from public sources that present information for the general public or parents who are interested in AST. We examine word frequency, bigrams, and concordances in these selected documents, and also identify key topics presented by each stakeholder group. We then compare the findings from these documents to a selection of studies on AST and explore the extent to which there is concordance between the scholarly literature on AST and materials shared with the public.

@@

Factors that influence AST have been presented and organized using a socio-ecological model (SEM) [@mitraIndependentMobilityMode2013] or systems model [@badlandDevelopmentSystemsModel2016] whereby children's travel behaviour is understood within the context of household, social, neighbourhood, and policy environments. The SEM helps to identify multiple determinants that need to be addressed by school-based interventions to increase AST, with the consensus being that interventions should target multiple levels [@mitraIndependentMobilityMode2013]. Rather than describing an extensive list of factors that influence AST and mode choice in Canada [e.g., @mammenUnderstandingDriveEscort2012; @mitraIndependentMobilityMode2013; @rothmanDeclineActiveSchool2018; @wilsonUnderstandingChildParent2018], we discuss a few potentially modifiable determinants that may be targeted for change through STP.

Children's mode choice to school is strongly influenced by their parents' travel behaviours and the complexity of their household's travel needs [@buliungLivingJourneySchool2021], as well as daily parental support [@mahDoesParentalSupport2017a]. This indicates that shifting parental perceptions and habits is important. Convenience and inclement weather have been cited by parents as barriers to AST [@buliungSchoolTravelPlanning2011]. Parental perceptions of the built or school environment [@demeesterParentalPerceivedNeighborhood2014; @panterAttitudesSocialSupport2010] and their children's skills [@mammenUnderstandingDriveEscort2012; @faulknerWhatQuickestEasiest2010] also influence mode choice to school.

Distance between home and school is most negatively associated with AST [@ikedaAssociationsChildrenActive2018; @mammenUnderstandingDriveEscort2012; @pontEnvironmentalCorrelatesChildren2009; @rothmanDeclineActiveSchool2018] with less AST reported among children who have to travel farther to school. Many studies have also found that the quality of the built environment along the route to school and around the school site [@ikedaAssociationsChildrenActive2018; @rothmanActiveSchoolTransportation2021] and provision of active travel infrastructure [@chenPromotingActiveStudent2018; @pontEnvironmentalCorrelatesChildren2009] facilitate AST. Canadian youth report that they feel most safe bicycling on streets in their neighbourhood or that have low volumes of traffic [@TACbikeinfra2020]. Finally, concerns about traffic and strangers have been reported by parents who drive their children to school [@mammenUnderstandingDriveEscort2012], which highlights that the volume and speed of cars can be a concern or deterrent for AST.

School travel planning in Canada and relevance of framing analysis
======================

School travel planning (STP) has been implemented in Canada since at least the late 2000s. Within the STP process, facilitators establish multi-sector committees which intervene at the participating school through a range of activities consisting of _[E]ducation_ strategies, _[E]ncouragement_ through in-person events or programs, _[E]ngineering_ improvements to or around the school site, and _[E]nforcement_ of traffic speeds around schools [or 4Es, @langUnderstandingModalChoice2011; @mammenActiveSchoolTravel2014]. Following the first large-scale evaluation of STP as an intervention in Canada, AST rates increased and 14% of parents reported a perceived reduction in traffic outside the school [@buliungSchoolTravelPlanning2011]. Assessments of the efficacy of STP [@buttazzoniPromotingActiveSchool2019; @mammenActiveSchoolTravel2014] indicate its potential to encourage behaviour change and adoption of AST.

By providing a "central organizing idea or story line" [@gamsonChangingCulture1987] framing of AST seems particularly important to shift parental attitudes and perceptions, given their reported influence on children's travel mode to school. Scholars in the field of political communications have proposed that communicators, such as the media or institutions, construct the narrative of a frame for policy positions or public issues in order to activate or restrict a particular response in the intended audience [@panFramingAnalysisApproach1993]. Organized groups of stakeholders can employ similar methods to attract attention to particular issues. Framing can be used to position existing solutions as suitable to address particular issues [@mahFramecriticalPolicyAnalysis2014], which may prevent the public from being aware of other policy approaches that challenge the status quo. The way policy issues are framed is ultimately important to understand because it plays a role in either altering or preserving the existing social perceptions. Finally, the ideas that are communicated by a particular frame can be either positive or negative [@waygoodCO2ValenceFraming2018]. There is evidence that negative framing may be more effective at motivating individuals, for example, to change transport modes to reduce CO2 emissions [@waygoodCO2ValenceFraming2018].

Framing of AST could also affect broader support in the community. Municipal representatives are perceived to be instrumental but the involvement of other stakeholder groups (e.g., busing consortia representatives and local residents) can be lacking [@buttazzoniSupportingActiveSchool2018]. In Ontario, it is reasonable to say that AST has become a policy issue on the education and public health agendas, supported, albeit in a modest way, by financial contributions from the provincial government. Still, the support from a range of municipal representatives [see @buttazzoniSupportingActiveSchool2018; @mammenPuttingSchoolTravel2015] demonstrates that the AST issue is on the political agenda.

Since STP activities heavily focus on education or encouragement [@buliungSchoolTravelPlanning2011; @mammenActiveSchoolTravel2014; @buttazzoniSupportingActiveSchool2018], it is important that parents and communities are receptive to the goals of STP [@buttazzoniSupportingActiveSchool2018]. Parents have been found to express different understandings, language, and perceptions than planners of how the built environment can influence school travel [@buliungLivingJourneySchool2021]. This is also true when it comes to other factors like convenience of different modes to school [@langUnderstandingModalChoice2011]. STP stakeholders must pay special attention to parents' understanding of the decline of AST as a problem, which may affect their receptivity to proposed solutions.

The success of STP interventions would likely depend on parental judgments of factors that are related to school travel, as well as support from key policy makers. For example, parents have been found to view mixed land use as conducive for driving, despite transport planners viewing neighbourhoods with mixed uses as key for encouraging more active travel [@buliungLivingJourneySchool2021]. Therefore, stakeholders involved in STP must make conscious choices about the proposed solutions and potential benefits of AST that are communicated to parents and the general public through publicly available content to convey its importance as a policy issue and to facilitate adoption of AST.

@@

We assembled a collection of publicly available documents that were sourced online from the main stakeholder groups involved in STP initiatives in Ontario: i) school boards (public or Catholic and English-speaking only); ii) municipal governments; and iii) transportation consortia. The latter involve a collaboration between municipal regions and school boards to deliver more efficient and timely transportation services to schools in each region. Non-profit organizations, police services, and advocacy groups are other stakeholders who often play a role in supporting AST and/or STP, but this study does not include any documents from these groups because they do not consistently participate in STP initiatives in Ontario.

@@

Webpages from STP stakeholder groups were included in our analysis if they were findable. This primary criterion was important since our analysis pertains to how such issues are framed to the general public. Thus, we included only webpages that were readily accessible, which we defined as requiring no more than four links from the initial Google search to reach.

>>>>

---
title: "Framing active school travel in Ontario: , or how spinach is good for you"
author:
  - name: Elise Desjardins
    email: desjae@mcmaster.ca
    affiliation: McMaster University
  - name: Jason Lam
    email: lamj54@mcmaster.ca
    affiliation: McMaster University
  - name: Darcy Reynard 
    email: reynard@ualberta.ca
    affiliation: University of Alberta
  - name: Damian Collins
    email: damian.collins@ualberta.ca
    affiliation: University of Alberta
  - name: E. Owen D. Waygood
    email: owen.waygood@polymtl.ca
    affiliation: Polytechnique Montreal
  - name: Antonio Paez
    email: paezha@mcmaster.ca
    affiliation: McMaster University
    correspondingauthor: true
    footnote: 1
address:
  - code: McMaster University
    address: School of Earth, Environment and Society, 1280 Main St. W, Hamilton, Ontario, Canada L8S 4K1
  - code: University of Alberta
    address: Department of Earth and Atmospheric Sciences, 1-26 Earth Sciences Building, Edmonton, Alberta, Canada T6G 2E3
  - code: Polytechnique Montreal
    address: Department of Civil, Geological and Mining Engineering, 2500, chemin de Polytechnique Montreal, Montreal, Quebec, Canada H3T 1J4
footnote:
  - code: 1
    text: "Corresponding author."
abstract: |
  Active school travel (AST) is a relevant policy issue in Ontario, Canada, as reflected in provincial support for school travel planning (STP) efforts. Two pillars of school travel plans are education and encouragement. For these pillars to stand, all relevant stakeholders must be cognizant of and understand well the issues and stakes. For this reason, how AST is communicated to them is essential. In this research we adopt frame analysis to investigate the way various organizations communicate around AST to the public. A frame is a central organizing idea or story line that provides meaning to a particular phenomenon, and helps to set the parameters for conversations about policy. In the case of AST, framing can influence what policy alternatives are perceived as available by children, parents, and their wider communities. To investigate framing of AST in Ontario we use natural language processing techniques to analyze publicly available documents from Ontario stakeholders involved in school travel planning. We then compared the findings from these documents to a selection of academic studies on AST. We conclude that framing of AST in Ontario is mostly empirical-scientific in style, and largely in agreement with academic research on AST. On the other hand, it it does not appeal to moral reasons for change, and is therefore oddly soothing and lacking in urgency. Furthermore, the frames used download responsibility for change to households, which limits the scope of policy alternatives by keeping collective action out of the frame.

keywords: 
  - active school travel
  - school travel planning
  - natural language processing
  - framing analysis
journal: "Transportation Research Part A: Policy and Practice"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: bibliography.bib
linenumbers: false
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: natbib
---

@@

Walking and bicycling to school, commonly known as active school travel (AST), have declined in North America for decades [@rothmanDeclineActiveSchool2018], but the appetite for these modes is beginning to rise in many Canadian communities. For instance, research in Toronto, Ontario found that 40% of children who were driven to school would like to travel by bicycle instead [@laroucheRatherBikeSchool2016]. Another study in London, Ontario, reported similar findings with respect to children's preference for active travel [@larsenRouteBasedAnalysisCapture2012]. From a public policy perspective, increasing active travel to school represents a major opportunity to improve children's health and wellbeing [@faulknerActiveSchoolTransport2009], to mitigate traffic and improve perceptions of safety [@rothmanAssociationsParentsPerception2015], as well as to reduce carbon and other noxious emissions [@bearman2014modelling; @goodman2019scenarios].

To promote this trend, in 2017 the Government of Ontario began to support communities develop AST initiatives through a program called _Ontario Active School Travel_ (OAST). By 2021, the program, run by _Green Communities Canada_ had awarded over CAD 2 million and provided resources to over 25 projects across Ontario [@GreenCommunities2016]. A popular intervention in many communities is school travel planning (STP) [see @GreenCommunitiesProjects]. In this context, STPs are "school-specific" interventions led by a facilitator who works with a committee of stakeholders from diverse sectors including education, planning, transportation, and public health to develop action plans [@buliungSchoolTravelPlanning2011; @mammenSchoolTravelPlanning2014]. The five-step process involves identifying barriers to AST and implementing solutions or activities to make AST safer and more convenient: 1) setup of the program; 2) data collection and problem identification; 3) action planning; 4) implementation; 5) monitoring and evaluation [@buliungSchoolTravelPlanning2011; @langUnderstandingModalChoice2011].

How AST is presented to a target audience (e.g., parents or the general public) is essential to define the parameters of what is possible policy-wise. Drawing from concepts in communication studies, the objective of this paper is to investigate how AST is framed in Ontario. Gamson and Modigliani [-@gamsonChangingCulture1987] define a frame as a “central organizing idea or story line that provides meaning” to a particular phenomenon. A frame can enable individuals "to locate, perceive, identify, and label" information pertaining to various dimensions of an issue [@goffmanFrameAnalysisEssay1974]. From this perspective, stakeholders involved in STP efforts play an important role in shaping public perception about AST by generating frames that  create meaning, thus influencing the degree of attention and recognition of a "problem" that needs to be addressed through behaviour change or new policies.

Our research is supported by natural language processing techniques, applied to the analysis of documents produced by three STP stakeholder groups in Ontario (municipalities, school boards, and transportation consortia). We assemble a corpus of texts from public sources that present information for the general public or parents who are interested in AST and examine word frequency, bigrams, and concordances in these selected documents. Furthermore, we use topic analysis to identify themes presented by each stakeholder group. We then compare the findings from these documents to a selection of studies on AST and explore the extent to which there is concordance between the scholarly literature on AST and materials shared with the public. The results of this analysis, along with spot checks of the corpus by the authors, lead us to conclude that AST if framed as a matter of individual choices instead of collective action. The message is reasonable but anodyne, and it fails to communicate a sense of urgency about health and environmental issues, which may help to explain the limited success of policy.

<!-- Compare the definition of "frame" to that of "narrative":

FRAME
A frame is a central organizing idea or story line that provides meaning to a particular phenomenon, and that enables individuals to locate, perceive, identify, and label information pertaining to various dimensions of an issue.

NARRATIVE (See [@liu2023narrative]

A narrative is defined as “a representation of connected events and characters that has an identifiable structure, is bounded in space and time, and contains implicit or explicit messages about the topic being
addressed”)
-->

@@

<!--Factors that influence AST have been presented and organized using a socio-ecological model (SEM) [@mitraIndependentMobilityMode2013] or systems model [@badlandDevelopmentSystemsModel2016] whereby children's travel behaviour is understood within the context of household, social, neighbourhood, and policy environments. The SEM helps to identify multiple determinants that need to be addressed by school-based interventions to increase AST, with the consensus being that interventions should target multiple levels [@mitraIndependentMobilityMode2013]. -->

Rather than describing an extensive list of factors that influence AST and mode choice for travel to school in Canada [e.g., @mammenUnderstandingDriveEscort2012; @mitraIndependentMobilityMode2013; @rothmanDeclineActiveSchool2018; @wilsonUnderstandingChildParent2018], we focus on two key aspects that may be targeted from a policy perspective: perceptions and behaviors by parents and built environmental factors.

<!-- Influence of perceptions -->

First, research has found that parents are key decision-makers with respect to children's form of travel to school [@rothmanAssociationsParentsPerception2015]. Particularly, parental perceptions of the built or school environment [@demeesterParentalPerceivedNeighborhood2014; @panterAttitudesSocialSupport2010; @mandic2020differences] and their children's skills [@mammenUnderstandingDriveEscort2012; @faulknerWhatQuickestEasiest2010] tend to influence mode choice to school. Furthermore, children's mode choice to school is strongly influenced by their parents' travel behaviours, the complexity of their household's travel needs [@buliungLivingJourneySchool2021], as well as daily parental support [@mahDoesParentalSupport2017a]. This suggests that shifting parental perceptions and habits is important.

<!-- Physical factors -->

With respect to the built environment, distance between home and school is most negatively associated with AST [@ikedaAssociationsChildrenActive2018; @mammenUnderstandingDriveEscort2012; @pontEnvironmentalCorrelatesChildren2009; @rothmanDeclineActiveSchool2018] with less AST reported among children who have to travel farther to school. Many studies have also found that the quality of the built environment along the route to school and around the school site [@ikedaAssociationsChildrenActive2018; @rothmanActiveSchoolTransportation2021] and provision of active travel infrastructure [@chenPromotingActiveStudent2018; @pontEnvironmentalCorrelatesChildren2009] facilitate AST. Canadian youth report that they feel most safe bicycling on streets in their neighbourhood or that have low volumes of traffic [@TACbikeinfra2020]. Finally, concerns about traffic and strangers have been reported by parents who drive their children to school [@mammenUnderstandingDriveEscort2012]: ironically, concerns about traffic become a self-fulfilling prophecy.

School travel planning in Canada and relevance of frames {#stp-frames}
======================

AST has become a policy issue on the education and public health agendas in Ontario, supported, albeit in a modest way, by financial contributions from the provincial government. Still, the support from a range of municipal representatives [see @buttazzoniSupportingActiveSchool2018; @mammenPuttingSchoolTravel2015] demonstrates that the AST issue is on the political agenda. One of the main vehicles to promote AST is school travel planning (STP), which has been implemented in Canada since at least the late 2000s. Within the STP process, facilitators establish multi-sector committees which intervene at the participating school through a range of activities consisting of _[E]ducation_ strategies, _[E]ncouragement_ through in-person events or programs, _[E]ngineering_ improvements to or around the school site, and _[E]nforcement_ of traffic speeds around schools [or 4Es, @langUnderstandingModalChoice2011; @mammenActiveSchoolTravel2014]. Following the first large-scale evaluation of STP as an intervention in Canada, AST rates increased and 14% of parents reported a perceived reduction in traffic outside the school [@buliungSchoolTravelPlanning2011]. Assessments of the efficacy of STP [@buttazzoniPromotingActiveSchool2019; @mammenActiveSchoolTravel2014] indicate its potential to encourage behaviour change and adoption of AST.

The relevance of policy frames is highlighted by @beland2014developing, who remarks "framing processes as public relations are central to policy issues related to sustainable transportation". Seeing how highly influential parental perceptions are to AST, it is important to understand AST-related communications, since the construction of frames for policy positions or public issues can activate or restrict a particular response in the intended audience [@panFramingAnalysisApproach1993]. Besides parents and children, framing can also affect broader policy support for AST in the community. Municipal representatives are perceived to be instrumental but the involvement of other stakeholder groups (e.g., busing consortia representatives and local residents) can be lacking [@buttazzoniSupportingActiveSchool2018]. In this sense, framing can be used to position existing solutions as suitable to address particular issues [@mahFramecriticalPolicyAnalysis2014], which may prevent the public from being aware of other policy approaches that challenge the status quo. Contrariwise, framing can be instrumental to revealing more diverse policy options than more diverse policy options than conventional wisdom would have [@bosomworth2015climate]. Ultimately, it is important to understand framing of policy issues because creating meaning through frames contributes to the constant process of shaping of our collective identities, uncovering what we collectively find (un)acceptable and why [@beland2014developing].

Given the heavy focus of STP activities on education or encouragement [@buliungSchoolTravelPlanning2011; @mammenActiveSchoolTravel2014; @buttazzoniSupportingActiveSchool2018], it is important for the issues to be effectively framed. In the past, parents have been found to express different understandings, language, and perceptions than planners of how the built environment can influence school travel: while parents tend to view mixed land use as conducive for driving, transport planners see mixed uses as a key for encouraging more active travel [@buliungLivingJourneySchool2021]. Similarly mismatched understandings have been detected with respect to factors like the convenience of different modes to school [@langUnderstandingModalChoice2011]. As a consequence, STP stakeholders must pay special attention to parents' and communities' understanding of the decline of AST as a problem, which may affect their receptivity to proposed solutions [@buttazzoniSupportingActiveSchool2018].

So framing is important, but not all frames are equally effective. Frames can appeal to economic reasons, to science and logic, or to moral obligation. For example, @severson2015moral report that overall support for climate change policy is increased more when scientific frames, secular moral frames, and economic equity frames are used, compared to economic efficiency or religious moral frames. Frames can also be positive or negative, and there is evidence that negative framing may be more effective at motivating individuals, for example, to change transport modes to reduce CO2 emissions [@waygoodCO2ValenceFraming2018]. As well, the effectiveness of the framing depends on the psychological distance to the relevant issues: how socially, temporally, geographically, or hypothetically close an individual perceives themselves to the issues can affect their support for various policies [@maiella2020psychological]. For this reason, stakeholders involved in STP must make conscious choices about how issues surrounding AST are communicated if the goal is to increase support for and buy-in of AST policies.

@@

We assembled a collection of publicly available documents that were sourced online from the main stakeholder groups involved in STP initiatives in Ontario: i) school boards (public or Catholic and English-speaking only); ii) municipal governments; and iii) transportation consortia. The latter involve collaborations between municipal regions and school boards to deliver more efficient and timely regional transportation services to schools. Non-profit organizations, police services, and advocacy groups are other stakeholders who often play a role in supporting AST and/or STP, but this study does not include any documents from these groups because they do not consistently participate in STP initiatives in Ontario.

@@

Webpages from STP stakeholder groups were included in our analysis if they were findable. This primary criterion was important since our analysis pertains communication of issues to the general public. Thus, we included only webpages that were readily accessible, which we defined as requiring no more than four links from the initial Google search to reach.
